---
title: "Day 1"
author: "Florian Gilberg"
date: "6 12 2021"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Day 1

```{r}
data <- read.delim("input.txt", sep="\n", header = FALSE)
```

This report indicates that, scanning outward from the submarine, the sonar sweep found depths of 199, 200, 208, 210, and so on.

The first order of business is to figure out how quickly the depth increases, just so you know what you're dealing with - you never know if the keys will get carried into deeper water by an ocean current or a fish or something.

To do this, count the number of times a depth measurement increases from the previous measurement. (There is no measurement before the first measurement.) In the example above, the changes are as follows:

199 (N/A - no previous measurement)
200 (increased)
208 (increased)
210 (increased)
200 (decreased)
207 (increased)
240 (increased)
269 (increased)
260 (decreased)
263 (increased)
In this example, there are 7 measurements that are larger than the previous measurement.

How many measurements are larger than the previous measurement?

## Part 1
```{r}
# iterate over each row, write if/else condition for each row, add as new text in column, count number of occurences

# Create new column
data <- tibble(data, "change" = character(1:length(data)))

for (i in 2:nrow(data)-1) { # dont to it for the first or last row
  
  # Subtract next row value from current row
  curr_value <- data$V1[i+1] - data$V1[i]
  
  # Use sign to check if value is -, + or 0
  sign_curr_value <- sign(curr_value)
  
  if (sign_curr_value == 1){ data$change[i+1] <- "increased"  }   # increase
  else if (sign_curr_value == -1) {data$change[i+1] <- "decreased" }   # decrease
  else if (sign_curr_value == 0){data$change[i+1] <- "unchanged" } # stayed the same
  else { break }
  
  
}


# Count number of increases / decreases

no <- table(data$change)
```
number of increases / decreases: `r no`

## Part 2

```{r}
# Create new column
data <- tibble(data, "sliding" = character(nrow(data)))

for (i in 2:nrow(data)-1) { # dont to it for the first or last row
  
  first_window <- data$V1[i] + data$V1[i+1] + data$V1[i+2]
  second_window <- data$V1[i+1] + data$V1[i+2] + data$V1[i+3]
  
  if (first_window < second_window){ data$sliding[i+1] <- "increased"  }   # increase
  else if (first_window > second_window) {data$sliding[i+1] <- "decreased" }   # decrease
  else if (first_window == second_window){data$sliding[i+1] <- "no change" } # stayed the same
  else { break }
  
}

table(data$sliding)
```

